---
Title: "METAHIT city-regions: estimates of disease case fatality"
author: "Chris Jackson <chris.jackson@mrc-bsu.cam.ac.uk>"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{METAHIT city regions disease case fatality estimates}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r,echo=FALSE}
knitr::opts_chunk$set(fig.width=7, echo=FALSE, warning=FALSE, message=FALSE)
```


```{r,echo=FALSE}
load(file="cfres.rda")
library(dplyr)
library(ggplot2)
source("constants.R")
```

## Overview

* Aim is to estimate case fatality for several chronic diseases and city regions of England to inform the METAHIT transport and health impact model. 

* 9 city regions (Bristol, Liverpool, London, West Midlands, Sheffield, North East, Greater Manchester, Nottingham, Leeds)

* Diseases: six cancers (colorectal, liver, stomach, lung/trachea/bronchus, female breast, uterus), cardiovascular disease (general), ischaemic heart disease, ischaemic stroke, other stroke, dementia, COPD _[though these may change?]_

* Data from the Global Burden of Disease study on incidence, prevalence and mortality are used.

* Bayesian methods are used to estimate case fatality based on these data and some model assumptions.  The principles behind the methods are explained in the [main disbayes vignette](ihdbristol.html).   Further details are given in this document on specific model assumptions that are used in this application.


## Model assumptions

* We assume the data arise from a three-state continuous-time Markov model with states representing no disease, disease and death from the disease, and parameters representing the transition rate from no disease to disease (incidence) and from disease to death from the disease (case fatality).

* Incidence and case fatality are assumed to depend on age, gender and area.   There is not enough data to estimate case fatality independently for each combination of age, gender and area.  However we can produce usefully-precise estimates under plausible assumptions about how the categories are related to each other.

* For all diseases, we assume that case fatality is a smooth spline function of age, with constant case fatality for age 50 and below.  Further assumptions are used where necessary to obtain usefully-precise area-specific estimates for particular diseases. 

* Two alternative assumptions are used for the _dependence on areas_ (implemented for all diseases)

	1. Independent, area-specific estimates, based on the model explained in the  [main disbayes vignette](ihdbristol.html). 

	2. Hierarchical model.	Allows data from other areas to inform the estimates from a weakly-informed area.  Explained in the  [hierarchical model vignette](hier.html).

* Two alternative assumptions are assumed for the _dependence on ages_

	1. No further restrictions on the smooth spline function

	2. _Non-decreasing_ case fatality with age (used for trachea/bronchus/lung cancer) 

	3. An "unsmoothed" model has also been implemented, where case fatalities for each year of age are estimated independently of the data from adjacent ages.    While this model is not used to generate results for METAHIT, it is still useful to identify where there is insufficient information about case fatality at particular ages (generally due to low prevalence at younger ages), thus extra assumptions about model structure are required, e.g. borrowing information between ages.  Results of these models are not presented in this document,.  An illustration is given in the [main disbayes vignette](ihdbristol.html). 

*  _Dependence on gender_.  A further set of models is fitted where the effect of gender is the same for each area.  In other words, there is no interaction between gender and area.  While this restriction is not necessary to produce _reasonable_ estimates, the intention is to produce more _precise_ estimates by borrowing information between the data for men and women. 

* Software implementation.  All models are implemented in Stan, as described in the [main disbayes vignette](ihdbristol.html).  Though the disbayes package currently only includes the independent area-specific model (1) - the hierarchical models could be included, but I wouldn't recommend them for non-expert use until they have been tested in other examples.



## Summary of results 

* Estimates of case fatality (posterior distributions) for all diseases, areas, and genders are given in "Full results plots" below.  A few additional plots are given to highlight specific aspects of these results, how these methods work and what is gained from the different modelling assumptions.

* For all cardiovascular diseases, good estimates are produced for all areas.  Those from the hierarchical model are generally more precise than from the independent-areas model.  These estimates also agree with estimates from Dismod II -- see the example in the _main vignette_

* Between-area variabiliy in case fatality for a specific age is shown in Figure 1.  This compares estimates from the hierarchical and independent-areas models of case fatality for IHD at age 60 for all areas.   Observe how the credible intervals from the hierarchical model are narrower and they have been _shrunk_ towards the mean over areas.  Informally, the estimates from each area are a weighted average of that area's data and the data from other areas. 

* Between-area variability in the curves of case fatality by age is shown in Figure 2.  This shows estimates from the hierarchical model of the between-area variability in the posterior median curve of case fatality by age.

	* _Discussion point_:  the uncertainty intervals overlap (see Figure 1), so the differences may not be seen as "significant". Though it's implausible that areas have exactly the same epidemiology - more likely that there are small differences, that may or may not be practically significant.  Our purpose is to inform a general health impact model (that will be used for many different purposes) with the best available data under the most realistic assumptions, while accounting for uncertainty, so I don't think there is a need to constrain area-specific estimates to be exactly the same. 

* For liver and stomach cancers (see "Full results plots") the area-specific data are too weak to provide meaningful estimates of case fatality.   This is shown by the independent-areas model producing implausibly wide interval estimates that decrease with age - essentially, these estimates are dominated by the vague priors used for the parameters governing the curve.  Simply borrowing information _between ages_ through the spline function is insufficient.  Plausible estimates can be obtained in these cases by borrowing information _between areas_ through the hierarchical model.   This is not a much stronger assumption - see the _hierarchical vignette_ for explanation - essentially we assume the area-level parameters are related through some unknown between-area distribution that we learn from the data.

* For trachea, bronchus and lung cancer, the models where the dependence of case fatality on age is an unrestricted spline function (not shown) still gives implausible estimates.  The estimates can be stabilised by assuming that case fatality is non-decreasing with age, under both the hierarchical and independent-areas models.   See "Full results plots" - the resulting estimates are roughly constant with age.    Figure 3 shows related ONS data on lung cancer survival in England by broad age groups - which verifies that the non-decreasing assumption is plausible.   Though note that this national data shows an increasing trend, while the model based on the GBD data appears to conclude that any evidence of variations with age are too weak, giving a constant estimate over all ages. 

* Figure 4 compares estimates from the models with and without the assumption that the effect of gender is the same in each area.  A single disease and area is shown (Bristol, COPD) but this is typical of the results in other settings, in that the two models give identical estimates where there is a lot of information, and slightly different but overlapping estimates where there is weak information, with the more restrictive model giving more precise estimates.   Statistical model comparison methods (e.g. cross-validation) might be used to determine whether the stronger assumption is adequate to describe the data. 


### Figure 1: between area variability in point and interval estimates, IHD, age 60 

```{r}
example_age <- 60
p <- cfres %>%
  filter(age == example_age,
         area %in% c(cityregions,regions_exc,regions_inc),
           disease %in% c("ishd") , 
           model %in% c("indep_areas", "hierarchical")) %>%
ggplot(aes(x=med, y=area, col=model)) +
    coord_cartesian(xlim=c(0.005, 0.03)) +
    facet_grid(rows=vars(disease), cols=vars(gender)) + 
    geom_point(position=ggstance::position_dodgev(height=0.4)) +
    geom_errorbarh(aes(xmin=lower95, xmax=upper95),
                   position=ggstance::position_dodgev(height=0.4)) +
    xlab(sprintf("Case fatality for age %s", example_age) )
p
```


### Figure 2: between area variability in age curves (IHD, point estimates) 

```{r}

p <- cfres %>%
  filter(disease %in% c("ishd") &
         area %in% cityregions, 
         model == "hierarchical") %>% 
ggplot(aes(x=age, col=area)) +
    geom_line(aes(y=med)) + 
    facet_grid(rows=vars(disease), cols=vars(gender)) +  
    coord_cartesian(ylim=c(0, 0.25), xlim=c(50,100)) +
    scale_x_continuous(breaks=seq(0,100,10)) + 
    ylab("Case fatality (median, 95% CrI)")  +
    xlab("Age (years)") 
p

```


### Figure 3: trachea, bronchus and lung cancer, male, point estimates from the model compared with related national estimates

Solid blue: estimates of net survival (converted from 5-year survival probabilities to rates) from lung cancer in England, by broad age groups, for adults diagnosed between 2013 and 2017 and followed up to 2018 (ONS).

```{r}

cdat <- readRDS(file="england_cancer_surv.rds") %>% 
  filter(`Cancer site` == "Lung", Sex == "Men") %>%
  select(agel, rate) 
cdat <- cdat %>%
    full_join(data.frame(agel=74, rate=cdat$rate[cdat$agel == 65]))

p <- cfres %>%
  filter(disease %in% c("tbalc") &
         area %in% cityregions,
         gender == "male" &
         model == "hierarchical") %>% 
  ggplot(aes(x=age, col=area)) +
  geom_line(aes(y=med)) + 
  facet_grid(rows=vars(disease), cols=vars(gender)) +  
  coord_cartesian(ylim=c(0, 0.5), xlim=c(15,100)) +
  scale_x_continuous(breaks=seq(0,100,10)) + 
  geom_step(aes(x=agel, y=rate), data=cdat, col="blue", size=2) + 
  ylab("Case fatality (median, 95% CrI)")  +
  xlab("Age (years)") 
p


```


### Figure 4:  assumptions about the gender effect compared 

Case fatality for COPD in Bristol. 

`hier_gender` indicates the model that assumes the gender effect is the same in each area.  `hierarchical` indicates the model with a different gender effect per area. 

```{r}
p <- cfres %>%
    filter(disease == "copd" &
           area == "bristol" &
           model %in% c("hierarchical", "hier_gender")) %>% 
ggplot(aes(x=age, col=model)) +
    geom_pointrange(aes(y=med, ymin=lower95, ymax=upper95),
                    position = position_dodge(0.2)) + 
    facet_grid(rows=vars(disease), cols=vars(gender)) +  
    coord_cartesian(ylim=c(0, 0.1), xlim=c(50, 100)) +
    scale_x_continuous(breaks=seq(0,100,10)) + 
    ylab("Case fatality (median, 95% CrI)")  +
    xlab("Age (years)") +
    ggtitle("Bristol")
p
```





## Full results plots

Estimates presented as two figures for each city: first figure including diseases other than cancer

* adaod: Alzheimer's disease and other dementia

* copd: chronic obstructive pulmonary disease

* crdd: cardiovascular diseases (general)

* iscs: ischaemic stroke

* ishd: ischaemic heart disease

* strk: stroke

_[ may also be type II diabetes here currently, but I think this will be treated differently ]_

and second figure including cancers

* brsc: breast cancer

* carc: colon and rectum cancer

* lvrc: liver cancer

* stmc: stomach cancer

* tbalc: trachea, bronchus and lung cancer

* utrc: uterine cancer 



```{r,cache=TRUE}

plot_area_diseases <- function(areaname, disease_set, yl=0.5){ 
    p <- cfres %>%
      filter(disease %in% disease_set &
             area == areaname &
             (model %in% c("indep_areas","hierarchical"))) %>% 
      ggplot(aes(x=age, col=model)) +
      geom_pointrange(aes(y=med, ymin=lower95, ymax=upper95),
                      position = position_dodge(0.2)) + 
      facet_grid(rows=vars(disease), cols=vars(gender)) +  
      coord_cartesian(ylim=c(0, yl)) +
      scale_x_continuous(breaks=seq(0,100,10)) + 
      scale_y_continuous(breaks=c(0,0.2,0.4)) + 
      ylab("Case fatality (median, 95% CrI)")  +
      xlab("Age (years)") +
      ggtitle(areaname)
    p
}
for (city in cityregions){
    print(plot_area_diseases(city, diseases_noncancer))
    print(plot_area_diseases(city, diseases_cancer))
}

```

## TODO

* Statistical model comparison (cross-validation / DIC-like method) for some of these

* Finish with final METAHIT diseases / settings.

* Discuss any other sources of uncertainty that are relevant to health impacts - e.g. trends through calendar time?



## Selling points for an applied statistics journal 

* A widely-used class of health impact simulation models requires disease incidence and case fatality to be represented as an explicit stochastic processes that we can simulate from (either at an aggregate or individual level) .

  - contrast this class of models with e.g. "comparative risk assessment" models, that are even more widely used to assess health impacts, but make stronger assumptions about age dependence

* Avalilability of data to inform different parts of that process varies with context, but typically the case fatality is the least well informed.  GBD is the most convenient source for models that can be adapted easily to different global contexts, but GBD don't publish estimates of case fatality.

* Barendregt et al method implemented in the DisMod II software is typically used in practice for this.  We can easily explain advantages of a Bayesian alternative (see [main disbayes vignette](ihdbristol.html)).

* Flaxman et al book describes the principles of Bayesian estimation of disease epidemiology from incomplete data.  We are applying those principles to a specific situation, and using a specific model, that was not described in that book, and that the associated software doesn't handle. 

* For an applied statistics journal, highlight interesting data features and modelling challenges that aren't explicitly discussed by Flaxman et al., e.g. 

  - borrowing info between ages, genders and areas, and statistical comparison of different assumptions 

  - more flexible and parsimonious spline models for age with smoothness controlled by penalisation 

  - nice way of constraining splines to be increasing where needed

  - hierarchical splines anchored on linear trends

* We develop an R package to implement our methods. 

* Remember the work is motivated by estimating health impacts of interventions/scenarios.  While the example is based on METAHIT transport model, this paper isn't about a specific health impact, rather to show a generic method that is applicable to different settings.   Emphasise that the method enables transparent uncertainty propagation for health impact estimates.
