---
output: html_document
editor_options: 
  chunk_output_type: console
---
Keep this nice and tidy.

```{r}

setwd("~/work/chronic/disbayes/metahit")
source("paper_analyses_header.r")
source("read_model_results.r")

```






# Data summary 

Raw point estimates of mort, inc and prev.   Just women for now.  Curves compared between areas, faceted by disease.   Shows diseases where prevalence and incidence are very low.  Mild differences between areas.    

```{r,eval=FALSE}

foo <- function(x){
  cat(deparse(substitute(x)))
}

ht <- 3
pdf("../../write/mort_data.pdf", height=ht)
gbdplot$areagender <- interaction(gbdplot$area, gbdplot$gender)
ggplot(gbdplot, aes(x=age, y=mort_prob, col=area, lty=gender, group=areagender)) + 
  facet_wrap(~disease, ncol=3, scales="free_y") + 
  coord_cartesian(xlim = c(50,100)) +
  ylab("Annual mortality probability") + 
  xlab("Year of age") +
  geom_ribbon(aes(ymin=mort_lower, ymax=mort_upper), fill="gray80", col="gray80") +
  geom_line() + 
  labs(col = "City region") +
  theme_bw() + 
  theme(strip.text.x = element_text(size = 7.5),
        legend.position = "none")
dev.off()

pdf("../../write/inc_data.pdf", height=ht)
ggplot(gbdplot, aes(x=age, y=inc_prob, col=area, lty=gender, group=areagender)) + 
  facet_wrap(~disease, ncol=3, scales="free_y") + 
  coord_cartesian(xlim = c(50,100)) +
  ylab("Annual incidence probability") + 
  xlab("Year of age") +
  geom_ribbon(aes(ymin=inc_lower, ymax=inc_upper), fill="gray80", col="gray80") +
  geom_line() + 
  labs(col = "City region") +
  theme_bw() + 
  theme(strip.text.x = element_text(size = 7.5),
        legend.position = "none")
dev.off()

pdf("../../write/prev_data.pdf", height=ht)
ggplot(gbdplot, aes(x=age, y=prev_prob, col=area, lty=gender, group=areagender)) + 
  facet_wrap(~disease, ncol=3, scales="free_y") + 
  coord_cartesian(xlim = c(50,100)) +
  ylab("Prevalence probability") + 
  xlab("Year of age") +
  geom_ribbon(aes(ymin=prev_lower, ymax=prev_upper), fill="gray80", col="gray80") +
  geom_line() + 
  labs(col = "City region", lty="") +
  theme_bw() + 
  theme(strip.text.x = element_text(size = 7.5),
        legend.spacing.y = unit(0.1, "cm"))
dev.off()


```



# Independent areas 

* Results from fitting disbayes model by disease, gender and area separately

* Stomach cancer and uterine cancer are done using national data. 

```{r,eval=FALSE}

## Convergence checks.   All OK with 5000 samples and the occasional model simplification
resnc <- resnh %>% filter(Rhat > 1.05)
table(resnc$disease, resnc$area, resnc$gender)


## Compare modal approx with posterior 
## Page of plots for a single area and gender.  Disease variation more important than area/gender variation  
## stmc, utrc use national point estimates 
## What are the current settings used for these? 

resall %>% 
  filter(var=="cf", disease=="Dementia", gender=="Female", 
         area=="Bristol", between(age, 70, 90), model=="Independent areas")

resoptunc <- readRDS("resoptunc_nonhier.rds") %>% 
  filter(!disease %in% c("Stomach cancer","Uterine cancer"))
resoptnat <- readRDS("resopt_nat.rds")
resoptunccf <- resoptunc %>%
  full_join(resoptnat) %>% 
  filter(var=="cf") %>%
  mutate(p50 = 1 - exp(-`par`),
         plow = 1 - exp(-`2.5%`),
         pupp = 1 - exp(-`97.5%`)) %>%
  filter(area=="Bristol" | area=="England", gender=="Female") %>%
  mutate(disease = fct_recode(disease, !!!disease_shorten),
           disease = factor(disease, levels = disease_order)) %>%
  filter(age %in% c(70, 75, 80, 85, 90)) %>%
  mutate(Algorithm = "Mode approximation")

pdf("../../write/opt_compare.pdf")

resp <- resall %>% 
  dplyr::filter(model %in% c("Independent areas","National"), var=="cf") %>%
  dplyr::filter(area=="Bristol" | (area=="England" & model=="National"), gender=="Female") %>%
  mutate(Algorithm="MCMC") %>%
  full_join(resoptunccf)
resp %>%
  filter(age %in% c(70, 75, 80, 85, 90)) %>%
  mutate(par = 1 - exp(-`50%`),
         plow = 1 - exp(-`2.5%`),
         pupp = 1 - exp(-`97.5%`)) %>%
  ggplot(aes(x=age, y=par, col=algorithm, lty=algorithm)) + 
  geom_errorbar(aes(ymin=plow, ymax=pupp), width=1.5,
              position=position_dodge(width=2), lwd=1.3) +
  geom_point(position=position_dodge(width=2), size=1.3) + 
  facet_wrap(vars(disease), ncol=2, scales="free_y") + 
  xlab("Age (years)") + ylab("Case fatality probability")

dev.off()



## Age curves for all diseases and areas 

pdf("../../write/all_cf_indep.pdf",width=6,height=6)

gg <- resall %>% 
  dplyr::filter(model %in% c("Independent areas","National"), var=="cf") %>%
  dplyr::filter(! (disease=="Uterine cancer" & age > 90)) %>%
  mutate(p50 = 1 - exp(-`50%`),
         plow = 1 - exp(-`2.5%`),
         pupp = 1 - exp(-`97.5%`)) %>%
  ggplot(aes(x=age, y=p50, group=area)) + 
  geom_ribbon(aes(ymin=plow, ymax=pupp), 
              fill="deepskyblue") +
  geom_line() + 
  facet_grid(cols=vars(gender), rows=vars(disease), scales = "free_y", switch = "y") + 
  xlab("Age (years)") + 
  ylab("Annual case fatality probability") + 
  xlim(60,100) + 
  theme(strip.text.y.left = element_text(angle = 0))

grob <- ggplotGrob(gg)
# gtable::gtable_show_layout(grob)
idx <- which(grob$layout$name %in% c("panel-5-2","panel-9-2"))
for (i in idx) grob$grobs[[i]] <- grid::nullGrob()
grid::grid.draw(grob)

dev.off()

```




# Hierarchical model

```{r,cache=TRUE}

## UTRC doesn't converge, so we use the empirical Bayes results 

resh %>%
  filter(disease=="Stomach cancer", gender=="Male", var=="cf") %>%
  ggplot(aes(x=age, y=`50%`, col=area, group=area)) + 
  geom_line()

resh %>% filter(gender=="Female", disease=="Stomach cancer", 
                var=="cf", age>40, age<60) %>% select(age,`50%`,Rhat) 
resnc <- resh %>% filter(Rhat > 2) 
resnc <- resh %>% filter(Rhat > 1.05) 
table(resnc$disease, resnc$gender)

## Look at precision of CRC estimates as exemplar. 
## Effective sample size of 1000 after a few hours
resh %>% filter(gender=="Male", disease=="Ischemic heart disease", age==60, var=="cf") %>% head
resh %>% filter(gender=="Male", disease=="Colon and rectum cancer", age==60, var=="cf") %>% head


## Between area variability in point and interval ests, hier and nonhier models  
## Specific age and gender, compare between areas 
## Not used in paper 

example_age <- 90
dat <- resall %>% filter(age==example_age, var=="cf", 
                         !(disease %in% c("Stomach cancer","Uterine cancer")),
                         #      disease %in% c("Colon and rectum cancer"),
                         model %in% c("Independent areas","Hierarchical"))
ggplot(dat, aes(x=`50%`, y=area, col=model)) +
  #   coord_cartesian(xlim=c(0.01, 0.05)) +
  facet_grid(rows=vars(disease), cols=vars(gender)) + 
  geom_point(position=ggstance::position_dodgev(height=0.4)) +
  geom_errorbarh(aes(xmin=`2.5%`, xmax=`97.5%`),
                 position=ggstance::position_dodgev(height=0.4)) +
  xlab(sprintf("Case fatality for age %s", example_age) )


## Between area variability in age curves, point ests.  CARC.  Shrinkage for oldest
## Use this in paper 

pdf("../write/hier_crc.pdf", width=6, height=4)

resall %>% 
  filter(var=="cf", 
         !disease %in% c("Stomach cancer","Uterine cancer"),
         disease %in% c(
           #   "Dementia"
           "Colon and rectum cancer"
           #       "Lung cancer"
         ),
         `97.5%` < 10,
         area %in% cityregions, 
         model %in% c("Independent areas","Hierarchical")) %>%
  mutate(model = relevel(factor(model), "Independent areas")) %>%
  ggplot(aes(x=age, col=area, group=area)) +
  geom_ribbon(aes(ymin=`2.5%`, ymax=`97.5%`), fill="gray80", col="gray80") +
  geom_line(aes(y=`50%`)) + 
  coord_cartesian(ylim=c(0, 1), 
                  xlim=c(70,  100)) +
  scale_x_continuous(breaks=seq(0,100,10)) + 
  ylab("Case fatality rate")  +
  facet_grid(cols=vars(model), rows=vars(gender)) + 
  xlab("Age (years)") + 
  theme_bw() + labs(col="")

dev.off()

```




# Hierarchical model with gender effect independent of area

Fuck me the Rhats are fucked 
Theyve basically all broken . 
Must be some unused parameter or wev that has been left in to break it 
Use the optim models to debug it 
looks like the incidence related parameters
beta_inc not in tidy but can't tell 




```{r}

resg <- resg %>% filter(disease != "Stomach cancer") # only national estimates
# convergence check
summary(resg$Rhat)  # 
resnc <- resg %>% filter(Rhat > 2) 
resnc <- resg %>% filter(Rhat > 1.1) 
resnc <- resg %>% filter(Rhat > 1.01) 

resnc <- resg %>% filter(Rhat > 1.1, disease=="Ischemic heart disease") 
resg %>% filter(grepl("beta_inc",var))

table(resnc$disease)
table(resnc$disease, resnc$area)

## point est curves compared between areas 
rescf <- resg %>% filter(gender=="Male", var=="cf")
rescf <- resg %>% filter(gender=="Female", var=="cf")
ggplot(rescf, aes(x=age, col=area)) +
  geom_line(aes(y=`50%`)) + 
  coord_cartesian(ylim=c(0, 1), 
                  xlim=c(50,  100)) +
  scale_x_continuous(breaks=seq(0,100,10)) + 
  ylab("Case fatality (median, 95% CrI)")  +
  facet_wrap(~disease, ncol=2) + 
  xlab("Age (years)") 

## Show whether gender effect is common between areas 
## todo pick diseases to illustrate.  IHD, LC biggesc contrast 
## how to show for all ages. 

dat <- resall %>%
  filter(var=="cf",
         model %in% c("Hierarchical","Hierarchical joint gender"),
         disease %in% c("Ischemic heart disease", "Stroke", "Lung cancer", "Colon and rectum cancer", "Dementia", "COPD")) %>%
  filter(disease %in% c("Ischemic heart disease", "Lung cancer")) %>%
  mutate(model = fct_recode(model, 
                            "Different between areas" = "Hierarchical",
                            "Common effect" = "Hierarchical joint gender"))
example_age <- 70 

#DEBUGME SHOW RAW MORT DATA FOR LC FOR AREAS, SEE IF GENDER DIFFERENCE
gbd %>% filter(ageyr==70, 
               disease=="Tracheal, bronchus, and lung cancer", 
               area %in% c("Bristol","Liverpool","West Midlands (city region)")) %>%
  mutate(p_mort = mort_num/mort_denom, p_prev=prev_num/prev_denom, mpratio=p_mort/p_prev) %>%
  select(sex, area, mort_num, mort_denom, p_mort, p_prev, mpratio)
## YES: MORT/PREV AGREES WITH CF 

i <- 3
resi <- readRDS(sprintf("results_hier_gender/res%s.rds",i))
resi <- readRDS(sprintf("results_hier/res%s.rds",3))
resi <- readRDS(sprintf("results_hier/res%s.rds",10))

resi[[1]] %>% 
  filter(var=="cf", between(age, 68, 72), area==1)
resi[[1]] %>% 
  filter(var=="cf", between(age, 68, 72), area==1)

gbd %>% filter(ageyr==70, 
               disease=="Tracheal, bronchus, and lung cancer", 
               area %in% c("Bristol","Liverpool","West Midlands (city region)")) %>%
  group_by(area) %>%
  mutate(mort_num=sum(mort_num), mort_denom=sum(mort_denom), 
         prev_num=sum(prev_num), prev_denom=sum(prev_denom)) %>%
  mutate(p_mort = mort_num/mort_denom, p_prev=prev_num/prev_denom, mpratio=p_mort/p_prev) %>%
  select(sex, area, mort_num, mort_denom, p_mort, p_prev, mpratio)

pdf("../write/gender.pdf", width=6, height=4) # NOTE one for IHD, one for lung cancer 

dat %>% filter(age==example_age) %>%
ggplot(aes(x=`50%`, y=area, lty=model, col=gender)) +
  #    coord_cartesian(xlim=c(0.001, 0.2)) +
  facet_wrap(~disease, nrow=1, scales="free_x") + 
  geom_point(position=ggstance::position_dodgev(height=0.6), size=1.5) +
  geom_errorbarh(aes(xmin=`2.5%`, xmax=`97.5%`),
                 position=ggstance::position_dodgev(height=0.6),
                 lwd = 0.7) +
  xlab(sprintf("Case fatality for age %s", example_age) ) + ylab("") + 
  scale_linetype_manual(values = c("Common effect" = "solid",
                                   "Different between areas" = "dashed"
                                   )) +
  labs(col="", lty="Gender effect") +
  theme_bw() + 
  theme(legend.text = element_text(size=7)) +  
  guides(linetype = guide_legend(reverse = TRUE))

dev.off()

```


## Lookup table for all results

100 * 9 * 18 * 2 * 3 results for each of inc and CF 

TODO just save this data frame in the metahit directory. there's no standard format

TODO append the results from optimisation.  OK to have multiple results for one case 
or just save in different format, since one has mode 

```{r}
app_tab <- resall %>%
  filter(var %in% c("cf","inc","rem")) %>%
  select(disease, var, gender, age, area, model, `25%`, `50%`, `97.5%`) %>%
  arrange(disease, var, gender, age, area, model)
app_tab %>% head
```



# Cross validatory comparison 

Compare nonhierarchical with hierarchical model, for all diseases and outcomes.  

Colon and rectum cancer is highlighted in the paper as a case where the hierarchical model makes a difference to the results, for ages 80+.   The hierarchical model gives better predictions here. 

```{r}
looall <- loonh %>% mutate(model="Independent areas") %>%
  full_join(looh %>% mutate(model="Hierarchical")) %>%
  filter(!(model=="National")) %>%
  select(age, gender, elpd_loo, model, area, disease, var) %>%
  group_by(model, gender, disease, var)

looall %>% 
  summarise(eisum = sum(elpd_loo)) %>%
  pivot_wider(names_from=model, values_from=eisum) %>%
  mutate(elpd_diff = `Hierarchical` - `Independent areas`)

looall %>% 
  filter(age > 80) %>%
  summarise(eisum = sum(elpd_loo)) %>%
  pivot_wider(names_from=model, values_from=eisum) %>%
  mutate(elpd_diff = `Hierarchical` - `Independent areas`)%>%
  filter(disease=="Colon and rectum cancer")
```

Graph of how predictive ability depends on age and area, for selected disease and outcome.  Hierarchical model is better for older ages.  Age 50 data seem to be strong enough that shrinkage is not needed.  Note that age 50 parameters are also assumed to apply to younger ages. 

```{r}
loosub <- looall %>% 
  filter(disease=="Colon and rectum cancer", gender=="Male", var=="mort") %>%
  ## sum ELPD by five year age group
  pivot_wider(names_from=model, values_from=elpd_loo) %>%
  mutate(elpd_diff = `Hierarchical` - `Independent areas`) %>%
  mutate(agegroup = cut(age, seq(0,100,5), right=FALSE, include.lowest = TRUE)) %>%
  group_by(area, agegroup, gender, var) %>%
  summarise(elpd_diff = sum(elpd_diff), .groups=c("drop")) %>%
  mutate(agemid = as.numeric(as.character(factor(agegroup, labels = seq(2.5, 97.5, by=5)))))

ggplot(loosub, aes(x=agemid, y=elpd_diff, colour=area)) +
  geom_hline(yintercept=0, col="gray", lwd=2) +
  geom_line() + 
  geom_point() + 
  scale_x_continuous(limits=c(30,100), breaks = seq(0,100,10)) + 
  geom_text(data=data.frame(x=c(50,50), y=c(-Inf,Inf)), 
            aes(x=x, y=y), 
            inherit.aes=FALSE,
            label=c("Independent areas preferred","Hierarchical preferred"), 
            vjust=c(0,1), hjust=c(0,0)) +
  directlabels::geom_dl(aes(label=area), 
                        method=list(directlabels::dl.combine("first.points", "last.points")))

```

Then compare hier with hier additive gender


```{r}

# elpd : high is good. positive elpd diff favours additive gender model 
loohg <- looall %>% 
  filter(model %in% c("Hierarchical","Hierarchical joint gender")) %>%
  select(var, area, age, gender, disease, model, elpd_loo) %>%
  pivot_wider(names_from=model, values_from=elpd_loo) %>%
  mutate(elpd_diff = `Hierarchical joint gender` - `Hierarchical`) %>%
  mutate(agegroup = cut(age, seq(0,100,5), right=FALSE, include.lowest = TRUE)) %>%
  group_by(var, area, gender, disease, agegroup) %>%
  summarise(elpd_diff = sum(elpd_diff), .groups=c("drop")) %>%
  mutate(agemid = as.numeric(as.character(factor(agegroup, labels = seq(2.5, 97.5, by=5)))))

loohg %>% 
  filter(disease %in% c("Ischemic heart disease","Lung cancer")) %>% 
  group_by(agemid, disease) %>%
  summarise(elpd = sum(elpd_diff)) %>%
ggplot(aes(x=agemid, y=elpd, col=disease, group=disease)) + 
  geom_hline(yintercept=0) + geom_line() + geom_point() 

# ELPD all over the place for different ages 
# positive: common effects.  negative: different effects 
# Different effects for lung cancer and dem
# not transparent though. 
# IHD vs LC is still biggest contrast 
# not sure makes sense to split by age.  model defines risk as smooth function of age

loohg %>%
  group_by(var, disease) %>%
  summarise(elpd = sum(elpd_diff)) %>%
  arrange(disease) %>%
  pivot_wider(names_from=var, values_from=elpd) %>%
  filter(!is.na(inc)) %>%
  mutate(rem=ifelse(is.na(rem), 0, rem),
         overall = inc+mort+prev+rem)


```






# Time trends 

Damn, Absolute incidence estimates are way off both the model results, 
but closest to the non trend one where they match up to age 70. 

why.  We are adjusting them to be consistent with prevalence and mortality data. 
But we want to give more weight somehow to current incidence data.  Why tho, if we also believe the current 
prev and mort data? are we trying to make them all fit into a bad model? 

Right the fitted values for year 2010 and earlier match up with current data.  There's a big jump around 2010 

But then they will carry on changing through time in the future if we use them in a model


```{r}

pdf("../write/trend_res.pdf")

datt <- gbd %>% 
  filter(area=="Leeds", disease=="Ischemic heart disease") %>%
  mutate(inc_prob = inc_num / inc_denom, 
         inc_rate = -log(1-inc_prob), 
         var = "Incidence", 
         gender = sex)

resoy <- resall %>% filter(model=="Time trends", 
                            year == 2017,
                            area=="Leeds", 
                            disease=="Ischemic heart disease",
                            var %in% c("inc")) %>% mutate(var="Incidence")
                           
rest <- resall %>% filter((model == "Independent areas") | (model=="Time trends" & year==2017),
                          area=="Leeds", 
                          disease=="Ischemic heart disease",
                          var %in% c("inc","cf")
#                          gender=="Male"
) %>% full_join(resoy)

resopt <- readRDS(file="resopt_trend.rds") %>% 
  filter(disease=="Ischemic heart disease",area=="Leeds",
         var %in% c("inc","cf"), year==99) %>%
  mutate(var = fct_recode(var, "Incidence"="inc", "Case fatality"="cf"))

rest %>% 
  filter(between(age, 50, 90)) %>% 
  mutate(var = fct_recode(var, "Incidence"="inc", "Case fatality"="cf"),
         model = fct_recode(model, "No time trends" = "Independent areas")) %>%
ggplot(aes(x=age, y=`50%`, col=model, fill=model, 
           group=interaction(gender, model))) +
  coord_cartesian(ylim=c(0.0,0.04)) +
  geom_ribbon(aes(ymin=`2.5%`, ymax=`97.5%`), alpha=0.2) +
  facet_wrap(vars(gender, var), nrow=2, scales="free_y") +
  geom_line() + 
  xlab("Age") + ylab("") + 
  labs(col="", fill="") + 
  geom_line(data=datt, aes(x=ageyr, y=inc_rate), inherit.aes = FALSE) +
  geom_line(data=resopt, aes(x=age, y=par), inherit.aes=FALSE, col="blue")

## Point estimates are accurate, so we can use them to diagnose the bias 

#  geom_line(data=resoy, aes(x=age, y=`50%`, col=factor(year), group=year), lty=2, inherit.aes = FALSE)
dev.off()

## Experimental sensitivity analyses to different trends
## what was it? 2003-2010 continues. saved in restrs 
## where's the conclusion from this?  estimates vary a lot

rest_saved <- rest %>% 
  filter(between(age, 50, 90), model=="Time trends", 
         gender=="Male", var %in% c("inc","cf")) 
res_saved <- resnh %>% filter(between(age, 50, 90), model=="Independent areas", disease=="Ischemic heart disease", area=="Leeds", gender=="Male", var %in% c("inc","cf"))
  
restrs <- restrs %>% 
  mutate(trendin = ifelse(inctrend & cftrend, "both", 
  ifelse(inctrend, "inc", ifelse(cftrend, "cf", "neither")))) %>%
  mutate(trendin = ifelse(trsens=="zero", "neither", trendin))

dat <- restrs %>% 
  filter(var %in% c("inc","cf"), model=="smoothed", trsens %in% c("hi", "zero"))
dat$age <- rep(1:100, length.out=nrow(dat))
dat %>%
  filter(age < 80) %>%
ggplot(aes(x=age, y=`50%`, col=trendin)) + 
  geom_line(lwd=2) + 
  facet_grid(cols=vars(var)) + 
  geom_line(data=datt %>% filter(gender=="Male") %>% mutate(var=fct_recode(var,"inc"="Incidence")), 
            aes(x=ageyr, y=inc_rate), inherit.aes = FALSE) 

+ 
  geom_line(data=rest_saved, col="black") + 
  geom_line(data=res_saved, col="blue")

restrs %>% 
  filter(var %in% c("inc","cf")) %>%
  mutate(age = rep(1:100, length.out=n())) %>%
  filter(model=="smoothed", age==60, var=="cf", cftrend==TRUE, inctrend==TRUE) 


```

Right this is saying that it doesn't make much difference to CF, but big diff to inc as in Scar
were the original plots wrong.   OK fine just axis limit distortion. 

So the only problem left is disagreement with current empirical incidence 
dismod2 types would say the model is "forcing consistency" 
evidence of conflict between current inc and current CF  
Hm is this evidence that the gbd data are consistent with themselves under a model that ignores trends.  

Hey wot it does match.   Wot about original independent-areas results 

cf - higher with inc or both, lower with cf or neither 
inc - higher with cf only, then neither, then both, then lowest with inc. 

So dose response is OK 
but they interact in unpredictable ways 
and need to fix the 5 year tihng 
and reconcile the modest changes here with the big changes from the real data

Compare real with artificial trend data.  Where does the difference come from? 
Looks reasonably representative of CF to me.  Age 70 value is generous 
Incidence: trend data is more noisy.  Pick an even higher value that's on the high end of the data. 

```{r}
ism <- inctrends_hi %>% as.data.frame() %>% 
  mutate(age = 1:100) %>%
  pivot_longer(names_to = "year", values_to = "p2017", cols = 1:100) %>%
  mutate(year = as.numeric(year)) %>% 
  filter(age %in% c(40,60,80,100), year > 1970) %>% 
  mutate(age = factor(age), model = "sens")

ihdtrends %>% 
  filter(age %in% c(40,60,80,100), year > 1970, gender=="Male", outcome=="Incidence") %>% 
  mutate(age = factor(age), model = "data") %>%
  full_join(ism) %>%
  ggplot(aes(x=year, y=p2017, col=age, lty=model)) + geom_line()

```



Well that's odd. 

With inc trends on. 
There's much less of an effect than there was with the real data. 
Consistent dose response effect 
CF trend makes no difference to current CF ests , but affects current inc ests. Wot? 

With cf trends on
Only an effect on cf if there is inc trends.
Consistent dose response when there's an effect
Inc est: both inc and cf trends make a difference.  moderate differences

Is this a bug in the sens analysis setup??   cf trends are not affecting cf
And why are there still those stupid 5 year artefacts for incidence?  didn't i transfer the new data? 

OK think og it a new way 
TREND IN: 
inc but not cf : 
cf but not inc : reduced inc 
both inc and cf:     reduced inc 

inc highest: 
inctrend false, cftrend true
inctrend true, cftrend true 
cftrend false.   
inc lowest 

cf highest with 
inctrend true, cftrend false or true 
inctrend false 